EuPPollNet: A European database of plant-pollinator networks

### 2024-09-06 Manuscript/EuPPollNet_Ms.qmd
edits

Although Europe contains a much larger number of flowering plants than pollinator species (~5 to 1 ratio according to our extrapolation from checklists), the observed number of pollinator species in the database was almost double that of the plants. This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018]. While animal-centered sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants. In addition, we found that the accumulation curve of pollinators per plant species does not saturate, which indicates low redundancy of pollinators and that many are regionally “rare”. Rare pollinators are functionally important for plant species at large scales [@winfree2018; @simpson2022], highlighting the need to conduct further sampling events to observe these rare species and to effectively understand and protect plant-pollinator biodiversity.

>>>>

Although Europe contains a much larger number of flowering plants than pollinator species (~5 to 1 ratio according to our extrapolation from checklists), the observed number of pollinator species in the database was almost double that of the plants. This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018]. While animal-centered sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants. In addition, we found that the accumulation curve of pollinators per plant species does not saturate, which indicates low redundancy of pollinators and that many are regionally “rare”. Rare pollinators can be functionally important for plant species at the landscape level [@winfree2018; @simpson2022], highlighting the need to conduct further sampling events to identify these rare species across different regions and to effectively understand and protect plant-pollinator biodiversity.
